Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-01-20 Grassmannian_cells.tex
Section 3 edits

\begin{lemma}
  For $V\in\Gr(H_m)$, $m\ge1$, the representation $\Sigma_2(P_{m+1}^V)^\sigma$ is also truncated preprojective.
\end{lemma}
\begin{proof}
  Apply the functor $\Sigma_2(-)^\sigma$ to the sequence \eqref{eq:truncated preprojectives}.
\end{proof}

@@

For $V\in \Gr(H_m)$, $m\ge1$, any subspace $W\subset V$ gives rise to an exact sequence
\[\xymatrix{0 \ar[r] & P_m\otimes(V/W) \ar[r]^(.6){ev} & P_{m+1}^W \ar[r] & P_{m+1}^V \ar[r] & 0},\]
where the left hand morphism above is the natural evaluation morphism coming from Lemma~\ref{le:truncated homomorphisms}.
Each such sequence has the following almost-split property for proper subrepresentations of $P_{m+1}^V$.
\sayD{The next results need some serious clean-up, can they be proven for $K(n)$ without passing to the universal cover?}
\begin{lemma}
  For $V\in \Gr(H_m)$, $m\ge1$, any proper subrepresentation $X\subsetneq P_{m+1}^V$ can be written as a direct sum of preprojective representations $P_r$ with $1\le r\le m$.
  Moreover, the upper pullback sequence
  \[\xymatrix{0 \ar[r] & P_m\otimes(V/W) \ar@{=}[d]\ar[r] & Y \ar[r]\ar[d]\ar@{}[dr]|(.7){\lrcorner} & X \ar[d]\ar[r] & 0 \\
    0 \ar[r] & P_m\otimes(V/W) \ar[r] & P_{m+1}^W \ar[r] & P_{m+1}^V \ar[r] & 0}\]
  splits for every proper subspace $W\subsetneq V$.
\end{lemma}
\begin{proof}
  For the first claim, we proceed by induction on $m$.
  When $m=1$, the dimension vector of $P_{m+1}^V$ is $(\codim_{H_m} V,1)$.
  In particular, it is immediate that each proper subrepresentation of $P_{m+1}^V$ is isomorphic to $P_1^k$ for some $0\le k\le l$.

  By Corollary~\ref{cor:truncated preprojective isomorphism}, the first claim for $m\ge2$ and $\codim_{H_m} V = 1$ follows from the first claim for $m-1$.

  Now we show that the first claim implies the second.
  Consider $W\subsetneq V$ and suppose the first claim holds for $P_{m+1}^V$.
  Observe that $\Ext(P_r,P_m)=0$ for $1\le r\le m$ and so the upper pullback sequence above will split giving the second claim.

  We proceed by induction on $\codim_V W$.  
  When $\codim_V W=1$, we have an exact sequence
  \[\xymatrix{0 \ar[r] & P_m\otimes(V/W) \ar[r] & P_{m+1}^W \ar[r] & P_{m+1}^V \ar[r] & 0}\] 
  with both morphisms irreducible by Lemma~\ref{le:irreducible morphisms} and the claim immediately follows.

  Suppose $W\subsetneq V$ with $\codim_V W>1$ and choose an intermediate subspace $W\subsetneq T\subsetneq V$.
  Given $X\subsetneq P_{m+1}^V$, we have the pullback diagram
  \[\xymatrix{ Y \ar[d]\ar[r]\ar@{}[dr]|(.7){\lrcorner} & Y' \ar[r]\ar[d]\ar@{}[dr]|(.7){\lrcorner} & X \ar[d]\\
  P_{m+1}^W \ar[r] & P_{m+1}^T \ar[r] & P_{m+1}^V}\]
  in which all horizontal morphisms of the upper row are split epimorphisms by induction.
  The result now follows since the outer square is also a pullback.
\end{proof}
\begin{corollary}
  Consider $V,W\in \Gr(H_m)$, $m\ge1$, with $W\subset V$.
  Given any proper subrepresentation $X\subsetneq P_{m+1}^V$ and any subrepresentation $Z\subset P_m\otimes(V/W)$, there is a subrepresentation of $P_{m+1}^W$ isomorphic to $Z\oplus X$ which fits into a commutative diagram
  \[\xymatrix{0 \ar[r] & Z \ar[d]\ar[r] & Z\oplus X \ar[r]\ar[d] & X \ar[d]\ar[r] & 0 \\
    0 \ar[r] & P_m\otimes(V/W) \ar[r] & P_{m+1}^W \ar[r] & P_{m+1}^V \ar[r] & 0}\]
\end{corollary}

\begin{lemma}
  For $V\in \Gr(H_m)$, $m\ge1$, a subrepresentation $X\subset P_m$ fits into a commutative diagram
  \[\xymatrix{0 \ar[r] & X \ar[d]\ar[r]\ar@{}[dr]|(.3){\ulcorner} & Y \ar[d]\ar[r] & P_{m+1}^V \ar@{=}[d]\ar[r] & 0 \\
    0 \ar[r] & P_m\otimes(V/U) \ar[r] & P_{m+1}^U \ar[r] & P_{m+1}^V \ar[r] & 0}\]
  with $\dim(V/U)=1$ if and only if the lower pushout sequence below splits: 
  \[\xymatrix{ 0\ar[r]& P_m\otimes(V/U) \ar[r] \ar[d]\ar@{}[dr]|(.3){\ulcorner} & P_{m+1}^U \ar[r]\ar[d] & P_{m+1}^V \ar@{=}[d]\ar[r] & 0\\
  0 \ar[r] & P_m\otimes(V/U)/X \ar[r] & Z \ar[r] & P_{m+1}^V \ar[r] & 0}.\]
  This occurs (?) if and only if the composed morphism
  \[\xymatrix{X \ar[r] & P_m \ar[r] & P_m^{\bar{V}} \ar[r] & P_{m-1}^{\bar{\bar{V}}}}\]
  is nonzero.
\end{lemma}
\begin{proof}
  For a proper subrepresentation $X\subsetneq P_m$ we have the following commutative diagram
  \[\tag{$\dagger$}\xymatrix{ & 0 \ar[r]\ar[d] & \Hom(P_{m+1}^U,P_m/X) \ar[d]\\
    0 \ar[r]\ar[d] & \Hom(P_m,P_m) \ar[r]\ar[d] & \Hom(P_m,P_m/X) \ar[d]\\
    \Ext(P_{m+1}^V,X) \ar[r] & \Ext(P_{m+1}^V,P_m) \ar[r] & \Ext(P_{m+1}^V,P_m/X)}\]
  From the bottom row of the diagram ($\dagger$), we see that there is a pushout diagram
  \[\xymatrix{0 \ar[r] & X \ar[d]\ar[r]\ar@{}[dr]|(.3){\ulcorner} & Y \ar[d]\ar[r] & P_{m+1}^V \ar@{=}[d]\ar[r] & 0 \\
    0 \ar[r] & P_m \ar[r] & P_{m+1}^U \ar[r] & P_{m+1}^V \ar[r] & 0}\]
  if and only if the lower pushout sequence below splits: 
  \[\xymatrix{ 0\ar[r]& P_m \ar[r]^{\epsilon}\ar[d]_{\rho}\ar@{}[dr]|(.3){\ulcorner} & P_{m+1}^U \ar[r]\ar[d] & P_{m+1}^V \ar@{=}[d]\ar[r] & 0\\
  0 \ar[r] & P_m/X \ar[r] & Z \ar[r] & P_{m+1}^V \ar[r] & 0}.\]
  The right hand vertical sequence of the diagram ($\dagger$) then shows that this occurs if and only if the map $\rho:P_m\to P_m/X$ factors through the map $\epsilon:P_m\to P_{m+1}^U$.

@@

\begin{definition}Let ${\bf j}:=(\underline{j_1},\underline{j_2},\ldots,\underline{j_r})$ be a sequence with $\underline{j_1}\in I_{n-1}$, $\underline{j_2},\ldots,\underline{j_{r-1}}\in I_{n-2}$ and $\underline{j_r}\in I_s$ for $s\leq n-2$. Define $j_1^c$ as before.
 We say that ${\bf j}$ is admissible if 
$$ j_{k,i}\neq j_{k-1}^{c}\text{ for all }k=2,\ldots, r-1,\,i=1,\ldots,n-2\text{ and for }i=1,\ldots,s\text{ if }k=r$$
where $j_{k}^c\in\{1,\ldots,n\}$ for $2\leq k\leq r-1$ is recursively defined as the unique element satisfying $j_k^c\neq j_{k,i}$ for $i=1,\ldots,n-2$ and $j_k^c\neq j_{k-1}^c$. 
\end{definition}

>>>>

\begin{lemma}
  \label{le:reflected truncated preprojectives}
  For $V\in\Gr(H_m)$, $m\ge2$, the representation $\Sigma_1(P_{m+1}^V)^\sigma$ is also truncated preprojective.
\end{lemma}
\begin{proof}
  By Corollary~\ref{cor:indecomposable truncated} and Lemma~\ref{le:unique morphisms}, $P_{m+1}^V$ is indecomposable but not rigid.
  In particular, $P_{m+1}^V$ does not have a summand isomorphic to $S_1$.
  Thus, following Remark~\ref{rem:reflection recursion}, we may apply the functor $\Sigma_1(-)^\sigma$ to the sequence \eqref{eq:truncated preprojectives} to get the exact sequence
  \[0\longrightarrow P_{m-1}\otimes V\xrightarrow{\Sigma_1(ev_V)^\sigma} P_m\longrightarrow \Sigma_1(P_{m+1}^V)^\sigma\longrightarrow 0\]
  which gives the claim.
\end{proof}

For $V\in \Gr(H_m)$, $m\ge1$, any subspace $W\subset V$ gives rise to an exact sequence
\[\xymatrix{0 \ar[r] & P_m\otimes(V/W) \ar[r]^(.6){ev} & P_{m+1}^W \ar[r] & P_{m+1}^V \ar[r] & 0},\]
where the left hand morphism above is the natural evaluation morphism coming from Lemma~\ref{le:truncated homomorphisms}.
Each such sequence has the following almost-split property for proper subrepresentations of $P_{m+1}^V$.
\begin{lemma}
  For $V\in \Gr(H_m)$, $m\ge1$, any proper subrepresentation $X\subsetneq P_{m+1}^V$ can be written as a direct sum of preprojective representations $P_r$ with $1\le r\le m$.
  Moreover, the upper pullback sequence
  \[\xymatrix{0 \ar[r] & P_m\otimes(V/W) \ar@{=}[d]\ar[r] & Y \ar[r]\ar[d]\ar@{}[dr]|(.7){\lrcorner} & X \ar[d]\ar[r] & 0 \\
    0 \ar[r] & P_m\otimes(V/W) \ar[r] & P_{m+1}^W \ar[r] & P_{m+1}^V \ar[r] & 0}\]
  splits for every proper subspace $W\subsetneq V$.
\end{lemma}
\begin{proof}
  For the first claim, we proceed by induction on $m$.
  When $m=1$, the dimension vector of $P_{m+1}^V$ is $(\codim_{H_m} V,1)$.
  In particular, it is immediate that each proper subrepresentation of $P_{m+1}^V$ is isomorphic to $P_1^k$ for some $0\le k\le l$.

  For $m\ge2$, we observe by induction that Lemma~\ref{le:reflected truncated preprojectives} implies any subrepresentation of $P_{m+1}^V$ which is not isomorphic to $P_1$ must be a direct sum of preprojective representations $P_r$ with $2\le r\le m$.
  Indeed, each of these is obtained from a subrepresentation of $\Sigma_1(P_{m+1}^V)^\sigma$ by applying the functor $\Sigma_2(-)^\sigma$ and the claim follows from the recursions in Remark~\ref{rem:reflection recursion}.

  Now we show that the first claim implies the second.
  Consider $W\subsetneq V$ and suppose the first claim holds for $P_{m+1}^V$.
  Observe that $\Ext(P_r,P_m)=0$ for $1\le r\le m$ and, since $X$ is a direct sum of preprojectives, the upper pullback sequence above must split.
\end{proof}
\begin{corollary}
  Consider $V,W\in \Gr(H_m)$, $m\ge1$, with $W\subset V$.
  Given any proper subrepresentation $X\subsetneq P_{m+1}^V$ and any subrepresentation $Z\subset P_m\otimes(V/W)$, there is a subrepresentation of $P_{m+1}^W$ isomorphic to $Z\oplus X$ which fits into a commutative diagram
  \[\xymatrix{0 \ar[r] & Z \ar[d]\ar[r] & Z\oplus X \ar[r]\ar[d] & X \ar[d]\ar[r] & 0 \\
    0 \ar[r] & P_m\otimes(V/W) \ar[r] & P_{m+1}^W \ar[r] & P_{m+1}^V \ar[r] & 0}\]
\end{corollary}

@@

\sayD{The next result needs some serious clean-up, can they be proven for $K(n)$ without passing to the universal cover?}
\begin{lemma}
  For $V\in \Gr(H_m)$, $m\ge1$, a subrepresentation $X\subset P_m$ fits into a commutative diagram
  \[\xymatrix{0 \ar[r] & X \ar[d]\ar[r]\ar@{}[dr]|(.3){\ulcorner} & Y \ar[d]\ar[r] & P_{m+1}^V \ar@{=}[d]\ar[r] & 0 \\
    0 \ar[r] & P_m\otimes(V/U) \ar[r] & P_{m+1}^U \ar[r] & P_{m+1}^V \ar[r] & 0}\]
  with $\dim(V/U)=1$ if and only if the lower pushout sequence below splits: 
  \[\xymatrix{ 0\ar[r]& P_m\otimes(V/U) \ar[r] \ar[d]\ar@{}[dr]|(.3){\ulcorner} & P_{m+1}^U \ar[r]\ar[d] & P_{m+1}^V \ar@{=}[d]\ar[r] & 0\\
  0 \ar[r] & P_m\otimes(V/U)/X \ar[r] & Z \ar[r] & P_{m+1}^V \ar[r] & 0}.\]
  This occurs (?) if and only if the composed morphism
  \[\xymatrix{X \ar[r] & P_m \ar[r] & P_m^{\bar{V}} \ar[r] & P_{m-1}^{\bar{\bar{V}}}}\]
  is nonzero.
\end{lemma}
\begin{proof}
  For a proper subrepresentation $X\subsetneq P_m$ we have the following commutative diagram
  \[\tag{$\dagger$}\xymatrix{ & 0 \ar[r]\ar[d] & \Hom(P_{m+1}^U,P_m/X) \ar[d]\\
    0 \ar[r]\ar[d] & \Hom(P_m,P_m) \ar[r]\ar[d] & \Hom(P_m,P_m/X) \ar[d]\\
    \Ext(P_{m+1}^V,X) \ar[r] & \Ext(P_{m+1}^V,P_m) \ar[r] & \Ext(P_{m+1}^V,P_m/X)}\]
  From the bottom row of the diagram ($\dagger$), we see that there is a pushout diagram
  \[\xymatrix{0 \ar[r] & X \ar[d]\ar[r]\ar@{}[dr]|(.3){\ulcorner} & Y \ar[d]\ar[r] & P_{m+1}^V \ar@{=}[d]\ar[r] & 0 \\
    0 \ar[r] & P_m \ar[r] & P_{m+1}^U \ar[r] & P_{m+1}^V \ar[r] & 0}\]
  if and only if the lower pushout sequence below splits: 
  \[\xymatrix{ 0\ar[r]& P_m \ar[r]^{\epsilon}\ar[d]_{\rho}\ar@{}[dr]|(.3){\ulcorner} & P_{m+1}^U \ar[r]\ar[d] & P_{m+1}^V \ar@{=}[d]\ar[r] & 0\\
  0 \ar[r] & P_m/X \ar[r] & Z \ar[r] & P_{m+1}^V \ar[r] & 0}.\]
  The right hand vertical sequence of the diagram ($\dagger$) then shows that this occurs if and only if the map $\rho:P_m\to P_m/X$ factors through the map $\epsilon:P_m\to P_{m+1}^U$.

@@

\begin{definition}Let ${\bf j}:=(\underline{j_1},\underline{j_2},\ldots,\underline{j_r})$ be a sequence with $\underline{j_1}\in I_{n-1}$, $\underline{j_2},\ldots,\underline{j_{r-1}}\in I_{n-2}$ and $\underline{j_r}\in I_s$ with $s\leq n-2$. Define $j_1^c$ as before.
 We say that ${\bf j}$ is admissible if 
$$ j_{k,i}\neq j_{k-1}^{c}\text{ for all }k=2,\ldots, r-1,\,i=1,\ldots,n-2\text{ and for }i=1,\ldots,s\text{ if }k=r$$
where $j_{k}^c\in\{1,\ldots,n\}$ for $2\leq k\leq r-1$ is recursively defined as the unique element satisfying $j_k^c\neq j_{k,i}$ for $i=1,\ldots,n-2$ and $j_k^c\neq j_{k-1}^c$. 
\end{definition}
